---
title: "abtestkit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{abtestkit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(abtestkit)
```

## What this package does

abtestkit is a toolkit for planning and analyzing binary-outcome A/B tests.
It covers both frequentist and Bayesian approaches side by side, aimed at
someone who understands what an A/B test is but has never worked with
Bayesian methods.

There are two phases: planning (before the experiment) and analysis (after the 
experiment). This vignette walks through both using a real marketing dataset.

## The dataset

The data comes from a Kaggle marketing A/B test where 588,101 users were
exposed to either an advertisement or a public service announcement (PSA)
placeholder. The outcome is whether the user converted (made a purchase).

We split this into two pieces:

- `historical_data`: 1,000 PSA users treated as pre-experiment baseline data.
  This is what we "know" before the experiment starts.
- `experiment_data`: 10,000 users per arm sampled from the remaining data.
  This is the actual experiment.

```{r load-data}
data(historical_data)
data(experiment_data)
```

## Phase 1: Pre-experiment planning

Before running an experiment, you need to answer two questions: how many
users do I need, and what do I believe going in?

The frequentist side answers the first question with a sample size
calculation. The Bayesian side answers the second by turning your historical
data into a prior distribution.

`pre_experiment_comparison` handles both at once. You pass in your historical
data, the column with the binary outcome, and the minimum detectable effect
(the smallest change worth detecting). Here we set `mde = 0.01`, meaning we
want to detect a 1 percentage point lift.

```{r pre-experiment}
mde <- 0.01
outcome_column <- "converted"
group_column <- "test.group"
control_value <- "psa"
treatment_value <- "ad"

pre_experiment_data <- pre_experiment_comparison(
  mde = mde,
  data = historical_data,
  outcome_column_name = outcome_column
)
```

The function returns two things: the sample size needed per group
(frequentist) and a Beta prior (Bayesian).

```{r pre-experiment-results}
n <- pre_experiment_data$n
experiment_prior <- pre_experiment_data$experiment_prior

cat("Required sample size per group:", n, "\n")
cat("Prior alpha:", round(experiment_prior$alpha, 3), "\n")
cat("Prior beta:", round(experiment_prior$beta, 3), "\n")
cat("Prior method:", experiment_prior$method, "\n")
```

The sample size tells you how long to run the test. The prior encodes what
you knew about conversion before seeing any experiment data. Because our
historical conversion rate is low (well below 50%), the prior was fit using
quantile matching, which handles skewed distributions better than moment
matching.

## Phase 2: Post-experiment analysis

Now we have experiment results. `post_experiment_comparison` runs both
analyses on the same data. The frequentist side runs a two-proportion z-test.
The Bayesian side updates the prior with each group's observed conversions
and compares the resulting posterior distributions.

```{r post-experiment}
post_experiment_data <- post_experiment_comparison(
  data = experiment_data,
  outcome_column_name = outcome_column,
  group_column_name = group_column,
  control_value = control_value,
  treatment_value = treatment_value,
  n = n,
  experiment_prior = experiment_prior
)
```

## Results

`full_comparison` puts everything together: side-by-side visuals and a
plain-language interpretation of what both methods found.

```{r comparison, fig.width = 10, fig.height = 5}
full_comparison(
  freq_results = post_experiment_data$freq_results,
  bayes_results = post_experiment_data$bayes_results,
  posterior_control = post_experiment_data$posterior_control,
  posterior_treatment = post_experiment_data$posterior_treatment
)
```

The left plot shows the frequentist confidence interval for the difference
in conversion rates. If the interval excludes zero (the red dashed line),
the result is statistically significant.

The right plot shows the Bayesian posterior distribution of the lift. The
black line is the expected lift, the dashed blue lines mark the 95% credible
interval, and the red dashed line is zero. If most of the distribution sits
to the right of zero, the Bayesian analysis favors the treatment.

The printed interpretation below the plots explains what each method
concluded, whether they agree, and the difference between a confidence
interval and a credible interval.
