---
title: "abtestkit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{abtestkit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(abtestkit)
```

## What this package does

LINK TO GITHUB: https://github.com/mbsperry-school/abtestkit

abtestkit is a toolkit for planning and analyzing binary-outcome A/B tests. An 
A/B test is an experiment that is used to compare the results of a randomized 
experiment where subjects can get one of two experiences. It is frequently used
in software, where users are given two different experiences on a website and 
employees are trying to answer the question: "Which is better?" Any time a big 
company releases a software update, it can be safely assumed that they ran some
variation of an A/B test before to get the confidence needed to ship to their 
whole user base. Scientists and engineers are able to have confidence in the 
changes in the software by using sound statistical methods that measure an 
effect on a small group of users, that would generalize to the entire base of 
users. It covers both frequentist and Bayesian approaches side by side, aimed 
at someone who understands what an A/B test is but has never worked with
Bayesian methods.

Most existing tooling makes you decide between frequentist and bayesian 
approaches. Comparing the two means putting together different packages that 
have different argument names and processes, which is confusing and discourages comparison. However, both of these methods are useful for different audiences, 
and the working statistician needs their coworkers and boss to readily 
understand their work in addition to being able to explain to non-technical 
audiences. Everyone that takes an intro-level statistics class is familiar 
with a p-value, which is the foundation of frequentist statistics. But the 
translation of bayesian statistical methods to stakeholders is much more 
straightforward. With this package, you don't have to choose, and through 
inspecting the functions, you can better understand for yourself what the 
differences are between these schools of thought. There are two phases: 
planning (before the experiment) and analysis (after the 
experiment). This vignette walks through both using a real marketing dataset.

## The dataset

The data comes from a Kaggle marketing A/B test where 588,101 users were
exposed to either an advertisement or a public service announcement (PSA)
placeholder. The outcome is whether the user converted (made a purchase).

We split this into two pieces:

- `historical_data`: 1,000 PSA users treated as pre-experiment baseline data.
  This is what we "know" before the experiment starts.
- `experiment_data`: 10,000 users per arm sampled from the remaining data.
  This is the actual experiment.

```{r load-data}
data(historical_data)
data(experiment_data)
```

## Phase 1: Pre-experiment planning

Before running an experiment, you need to answer two questions: 
- how many users do I need?
- what do I believe going in?

The frequentist side answers the first question with a sample size
calculation. The Bayesian side answers the second by turning your historical
data into a prior distribution. `pre_experiment_comparison` handles both at 
once. You pass in your historical data, the column with the binary outcome, 
and the minimum detectable effect (the smallest change worth detecting). Here 
we set `mde = 0.01`, meaning we want to detect a 1 percentage point lift.

```{r pre-experiment}
pre_experiment_data <- pre_experiment_comparison(
  mde = 0.01,
  data = historical_data,
  outcome_column_name = "converted"
)
```

The function returns two things: the sample size needed per group
(frequentist) and a Beta prior (Bayesian).

```{r pre-experiment-results}
n <- pre_experiment_data$n
experiment_prior <- pre_experiment_data$experiment_prior

cat("Required sample size per group:", n, "\n")
cat("Prior alpha:", round(experiment_prior$alpha, 3), "\n")
cat("Prior beta:", round(experiment_prior$beta, 3), "\n")
cat("Prior method:", experiment_prior$method, "\n")
```

The sample size tells you how long to run the test. The prior encodes what
you knew about conversion before seeing any experiment data. Because our
historical conversion rate is low (below 50%), the prior was fit using
quantile matching, which handles skewed distributions better than moment
matching.

## Phase 2: Post-experiment analysis

Now we have experiment results. `post_experiment_comparison` runs both
analyses on the same data. The frequentist side runs a two-proportion z-test.
The Bayesian side updates the prior with each group's observed conversions
and compares the resulting posterior distributions.

```{r post-experiment}
post_experiment_data <- post_experiment_comparison(
  data = experiment_data,
  outcome_column_name = "converted",
  group_column_name = "test.group",
  control_value = "psa",
  treatment_value = "ad",
  n = n,
  experiment_prior = experiment_prior
)
```

## Results

`full_comparison` puts everything together with side-by-side visuals and a clear interpretation of what both methods found.

```{r comparison, fig.width = 10, fig.height = 5}
full_comparison(
  freq_results = post_experiment_data$freq_results,
  bayes_results = post_experiment_data$bayes_results,
  posterior_control = post_experiment_data$posterior_control,
  posterior_treatment = post_experiment_data$posterior_treatment
)
```

The left plot shows the frequentist confidence interval for the difference
in conversion rates. If the interval excludes zero (the red dashed line),
the result is statistically significant. The right plot shows the Bayesian 
posterior distribution of the lift. The black line is the expected lift, the 
dashed blue lines mark the 95% credible interval, and the red dashed line is 
zero. If most of the distribution sits to the right of zero, the Bayesian 
analysis favors the treatment. The printed interpretation below the plots 
explains what each methodb concluded, whether they agree, and the difference 
between a confidence interval and a credible interval.


## Conclusions

You should not have to choose a school of thought to
run one experiment. The frequentist result gives you the significance threshold
that a review process or a skeptical colleague will expect. The Bayesian result
gives you the probability that the treatment is better, which is the quantity a
non-technical stakeholder was asking about in the first place. Producing both
from one set of inputs means the comparison is easy.

If you are comfortable with p-values and have been meaning to find out what
Bayesian methods would add, running your next test through
`pre_experiment_comparison()` and `full_comparison()` is the cheapest possible
way to find out. You get the analysis you would have run anyway, plus a second
answer to check it against.
